---
title: "Project 3 JB Section"
author: "Jonathan Burns & Michael Robinson"
date: "2023-10-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages needed:

```{r load-packages, message=FALSE}
library(tidyverse)
library(stringr)
library(dplyr)
library(ggplot2)
library(knitr)
library(reshape2)
library(RMySQL)
library(wordcloud2)
library(scales)
library(gridExtra)
library(plotly)
```


##Database Connection
While I have the data loaded in a SQL database, we wanted to use the raw data from the GitHub to show some cleaning and tidying skills we learned during class.
```{r}
con <- dbConnect(odbc::odbc(), "jbdb", timeout = 10)

```


```{r}
skills <- read.csv("https://raw.githubusercontent.com/jonburns2454/Project-3-DATA607/main/ds_general_skills_revised.csv")

software <- read.csv("https://raw.githubusercontent.com/jonburns2454/Project-3-DATA607/main/Data%20Science%20Career%20Terms%20-%20ds%20software.csv")

glimpse(skills)
glimpse(software)

```

Preliminary Work:

```         
- Checking for missing data
- Cleaning
```

```{r}
sum(is.na(skills))## 0

sum(is.na(software))## 0 

software.cleaned <- software[, -c(6:12)] # Removing the nonessential data columns

software.cleaned <- software.cleaned[-c(38:42), ]

skills.cleaned <- skills[-c(16:30), ] # Removing data that interferes with the format
    
```

##Further tidying:

Some regex to separeate the characters into numeric data and add it into a new data frame

```{r}
LinkedIn <- as.numeric(gsub("[%\\,]", "", skills.cleaned$LinkedIn))
Indeed <- as.numeric(gsub("[%\\,]", "", skills.cleaned$Indeed))
SimplyHired <- as.numeric(gsub("[%\\,]", "", skills.cleaned$SimplyHired))
Monster <- as.numeric(gsub("[%\\,]", "", skills.cleaned$Monster))
Keyword <- (gsub("[%\\,]", "", skills.cleaned$Keyword))

skillsDF <- data.frame(Keyword, LinkedIn, Indeed, SimplyHired, Monster)


```

Doing the same for our software data: (.S denotes software data)

```{r}
LinkedIn.S <- as.numeric(gsub("[%\\,]", "", software.cleaned$LinkedIn))
Indeed.S <- as.numeric(gsub("[%\\,]", "", software.cleaned$Indeed))
SimplyHired.S <- as.numeric(gsub("[%\\,]", "", software.cleaned$SimplyHired))
Monster.S <- as.numeric(gsub("[%\\,]", "", software.cleaned$Monster))
Keyword.S <- (gsub("[%\\,]", "", software.cleaned$Keyword))

softwareDF <- data.frame(Keyword.S, LinkedIn.S, Indeed.S, SimplyHired.S, Monster.S)


```

Adding in percentage variables to better envision the data breakdown.

##Software:

```{r}
softwareDF <- softwareDF %>% 
    mutate(LinkedInFreq = (LinkedIn.S / sum(LinkedIn.S))*100) %>% 
    mutate(IndeedFreq = (Indeed.S / sum(Indeed.S))*100) %>% 
    mutate(SimplyHiredFreq = (SimplyHired.S / sum(SimplyHired.S))*100) %>% 
    mutate(MonsterFreq = (Monster.S / sum(Monster.S))*100)


```

##Skills:

```{r}
skillsDF <- skillsDF %>% 
    mutate(LinkedInFreq = (LinkedIn / sum(LinkedIn))*100) %>% 
    mutate(IndeedFreq = (Indeed / sum(Indeed))*100) %>% 
    mutate(SimplyHiredFreq = (SimplyHired / sum(SimplyHired))*100) %>% 
    mutate(MonsterFreq = (Monster / sum(Monster))*100)



```

**Transforming the data further for visualization**

```{r}
meltedSkills <- melt(skillsDF, id.vars = "Keyword")


meltedSoftware <- melt(softwareDF, id.vars = "Keyword.S")


```

**Bar Chart : Count**

```{r}
meltedSkills %>% 
    filter(variable == "LinkedIn" | variable == "Indeed" | variable == "SimplyHired" | variable == "Monster") %>% 
ggplot(aes(x = Keyword, y = value, fill = variable)) +
    geom_bar(stat = "identity", position = "dodge", width = 0.8) +
    labs(x = "Keywords", y = "Count") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_manual(values = c("LinkedIn" = "coral1", "Indeed" = "firebrick4", "SimplyHired" = "aquamarine4", "Monster" = "chartreuse4"))+
    ggtitle("Keyword Count on Major Job Search Platforms (Skills)")+
    coord_flip()
```

**Bar Chart : Percent** *Note: This is the same idea, but instead of filtering our count numbers*

```{r}
meltedSkills %>% 
    filter(variable == "LinkedInFreq" | variable == "IndeedFreq" | variable == "SimplyHiredFreq" | variable == "MonsterFreq") %>% 
ggplot(aes(x = Keyword, y = value, fill = variable)) +
    geom_bar(stat = "identity", position = "dodge") +
    labs(x = "Keywords", y = "Frequency") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_manual(values = c("LinkedInFreq" = "coral1", "IndeedFreq" = "firebrick4", "SimplyHiredFreq" = "aquamarine4", "MonsterFreq" = "chartreuse4"))+
    ggtitle("Keyword Frequency on Major Job Search Platforms (Skills)")+
    coord_flip()
```

##Software

**Bar Chart - Software**

```{r}
meltedSoftware %>% 
    filter(variable == "LinkedIn.S" | variable == "Indeed.S" | variable == "SimplyHired.S" | variable == "Monster.S") %>% 
ggplot(aes(x = Keyword.S, y = value, fill = variable)) +
    geom_bar(stat = "identity", position = "dodge", width = 0.7) +
    labs(x = "Keywords", y = "Count") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_manual(values = c("LinkedIn.S" = "coral1", "Indeed.S" = "firebrick4", "SimplyHired.S" = "aquamarine4", "Monster.S" = "chartreuse4"))+
    ggtitle("Keyword Count on Major Job Search Platforms (Software)")+
    coord_flip()
```

**Bar Chart - Software Frequency**
```{r}
plot1 <- meltedSoftware %>% 
    filter(variable == "LinkedInFreq" | variable == "IndeedFreq" | variable == "SimplyHiredFreq" | variable == "MonsterFreq") %>% 
    ggplot(aes(x = Keyword.S, y = value, fill = variable)) +
    geom_bar(stat = "identity", position = "dodge", width = 0.7) +
    labs(x = "Keywords", y = "Count") +
    theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
    scale_fill_manual(values = c("LinkedInFreq" = "coral1", "IndeedFreq" = "firebrick4", "SimplyHiredFreq" = "aquamarine4", "MonsterFreq" = "chartreuse4")) +
    ggtitle("Keyword Count on Major Job Search Platforms (Software)") +
    coord_flip()

# Convert ggplot objects to plotly objects
interactive_plot1 <- ggplotly(plot1)

interactive_plot1_zoom <- interactive_plot1 %>%
  layout(xaxis = list(title = "Keywords"), yaxis = list(title = "Count"))

# Make the plot zoom interactive
interactive_plot1_zoom <- interactive_plot1_zoom %>%
  config(scrollZoom = TRUE)

interactive_plot1_zoom
```




##Word Cloud Analysis
**Using Skills Data to start**

Skills:
```{r}

top_keywords_skills <- meltedSkills %>% 
    group_by(Keyword) %>% 
    summarise(Frequency_LI = sum(as.numeric(value))) %>% 
    head(15) #adjusting this so all 15 keywords are present 

wordcloud2(top_keywords_skills, size = 0.7,
           minRotation = -pi/3, maxRotation = pi/3, shuffle = T,
           shape = 'circle', rotateRatio = 0.3)


```

Software:

```{r}
top_keywords_software <- meltedSoftware %>% 
    group_by(Keyword.S) %>% 
    summarise(Frequency_SF = sum(as.numeric(value))) %>% 
    head(100) #adjusting this so all 15 keywords are present 

wordcloud2(top_keywords_software, size = 0.8,
           minRotation = -pi/4, maxRotation = pi/4, shuffle = T,
           shape = 'circle', rotateRatio = 0.3)
```
